---
title: Get antequated with SOmap
author: null
date: '2019-03-20'
slug: get-antequated-with-somap
categories: []
tags: []
description: ''
featured: ''
featuredalt: ''
featuredpath: ''
linktitle: ''
editor_options: 
  chunk_output_type: console
---


## Polar maps are still hot! 

Last time we [got this far](http://antequated.org/2017/06/14/polar-maps-now/) by creating a very simplistic polar map and discussing some of the difficulties in customizing and finishing it. Since then [Dale Maschette](https://github.com/Maschette) discussed [these problems](https://twitter.com/StephdeSilva/status/1017631191854297088) at [useR Brisbane 2018](https://user2018.r-project.org) and [stirred up](https://twitter.com/Dale_Masch/status/1017693776452546560) multiple [discussions on twitter](https://twitter.com/databsurdist/status/1107797978423558149) about the joys of polar maps. 

Behold SOmap. 


## SOmap

To install the [SOmap](https://github.com/AustralianAntarcticDivision/SOmap) package use

```{r install, eval=FALSE}
remotes::install_github("AustralianAntarcticDivision/SOmap")
```
and see the [package readme and documentation](https://australianantarcticdivision.github.io/SOmap/) for further details. 

Creating a polar map is as easy as the following. 

```{r polar-go}
library(SOmap)
pmap <- SOmap()
pmap
```

To add our own data to the map we can use the `SOplot()` function, by default this function is geared to adding data to our existing map. 

```{r polar-add}
pmap
longitudes <- seq(-180, 175, by = 5)
latitudes <- approx(c(-48, -88, -48), n = length(longitudes))$y
SOplot(longitudes, latitudes, pch = 1:25)
```

The great feature here is that we only needed our data, points in longitude and latitude and a single call to `SOplot()` - the data is automtically added to the plot in the right way. 

We can also add spatial types to the plot. 

To start afresh from `SOmap()` above we can `print()` the object. 

```{r other-types}
print(pmap)
SOplot(SOmap_data$seaice_oct)
SOplot(SOmap_data$seaice_feb, col = "firebrick", lwd = 2, lty = 2)

```

In the previous post we had these desired features. 

* a local, specific, extent (this is surprisingly difficult to get right)
* a "graticule", i.e. lines of constant longitude and latitude
* other data on the map, again this can be suprisingly complicated
* labels
* aesthetics and scales for the data on the map
* accoutrements (scale bars and north arrow and the fancy white-black margin real maps use)


The extent is controlled by the data we give `SOauto_map()`, this also adapts the map projection in use to best suit the data given. 

```{r automap}
grid <- expand.grid(x= seq(105, 165, by = 5), y = seq(-75, -38, by = 8))
amap <- SOauto_map(grid$x, grid$y)
```

By default, the input data are plotted as points and as joined lines. The lines don't make sense here, so turn them off. 

```{r no-lines}
(amap <- SOauto_map(grid$x, grid$y, input_lines = FALSE))
```

As well as our very contrived set of grid points, more realistic data works as well. 

```{r tracks}
track <- SOmap::SOmap_data$mirounga_leonina
trackmap <-  SOauto_map(track$lon, track$lat, pcol = "yellow")
trackmap
```

While `SOplot()` provides the arguments for the plot method in use, `SOmap()` and `SOauto_map()` provide special argumens that match their base equivalents for points and lines, i.e. `pcol` and `lcol` respectively. 

We see immediately that the seal heading west raced home somewhat more quickly than its easterly counterpart. 

```{r tracktime}
SOauto_map(track$lon, track$lat, pcol = colourvalues::colour_values(track$date))

library(dplyr)
track %>% group_by(id) %>% 
  summarize(start = min(date), end =max(date), west = min(lon), east = max(lon))
```


## Next steps

...